---
title: "class11"
author: "Kaito Tanaka (PID: A15574793"
output: github_document
---
## PDB Statistics

```{r}
db <- read.csv("Data Export Summary (3).csv", row.names=1)
head(db)
```

> Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy?

```{r}
methods.sums <- colSums(db)
round((methods.sums/methods.sums["Total"])*100,2)
```

87.55% for X-Ray and 7.36% for NMR

> Q2: What proportion of structures in the PDB are protein?
```{r}
round(db$Total/methods.sums["Total"]*100,2)
```

87.36% of the structures in the PDB are proteins.

> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1828 HIV-1 protease structures

## Visualizing the HIV-1 protease

> Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

We can only see just one atom per water molecule (the oxygen), because the Hydrogen atom is too small to be seen through x-ray crystallography

> Q5: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have (see note below)?
residue number MK1902:04

# Using Bio3d

First, load the library
```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

Then Extract sequence from file
```{r}
pdbseq(pdb)
```

Three-letter codes of the amino acid sequence
```{r}
aa123(pdbseq(pdb))
```

The ATOM records
```{r}
head(pdb$atom)
```
