---
title: "R Functions"
author: "Kaito Tanaka"
date: "10/24/2021"
output:
  pdf_document: default
  html_document: default
---
First, install the package "bio3d"

install.packages("bio3d")

Our given code is as shown below: 

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

>Homework question: How would you generalize the original code above to work with any set of input protein structures?

Let us name a new function, protein_drug_plot

The input will be the specification of the file that is to be read.
The function will create a plot from the PDB file that will analyze the protein drug interactions and compare the residues to the B factor. Hence, the output will be the plot. 

```{r}
protein_drug_plot <- function(x){ 
  #here, we read
  s <- read.pdb(x)
  
  #here, we trim
  s.chainA <- trim.pdb(s, chain="A", elety="CA")
  
  #here, we set s.b
  s.b <- s.chainA$atom$b
  
  #Finally, plot
  plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor")
}

protein_drug_plot("4AKE")
protein_drug_plot("1AKE")
protein_drug_plot("1E4Y")
```
Make the cluster dendogram:
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

