---
title: "Class_17.Rmd"
author: "Kaito Tanaka"
date: "11/29/2021"
output: pdf_document
---

## Getting Started

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
head(vax$as_of_date)
tail(vax$as_of_date)
```

> Q1. What column details the total number of people fully vaccinated?

The column "persons_fully_vaccinated" details the total number of people fully vaccinated. 

> Q2. What column details the Zip code tabulation area?

The column "zip_code_tabulation_area" details the zip code tabulation area. 

> Q3. What is the earliest date in this dataset?

The earliest date is 2021-01-05. 

> Q4. What is the latest date in this dataset?

The latest date is 2021-11-23.

Get a quick overview of the dataset
```{r}
#skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset?

```{r}
ncol(vax)
```
9 numeric columns. 

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```
8256 NA values are present in "persons_fully_vaccinated".

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
(sum( is.na(vax$persons_fully_vaccinated) ) /
NROW(vax$persons_fully_vaccinated)) *100
```
10.17% of the data is missing here

##Working with Dates

We will use the **lubridate** package to make life a lot easier when dealing with dates and times: 

```{r}
#install.packages("lubridate")
library(lubridate)
```

what is today's date?
```{r}
today()
```

```{r}
#Specify our format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
#Example of what we can now do easily
today() - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset?

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
7 days have passed since the last update of the dataset.

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```
There are 47 unique dates in the dataset.

```{r}
#install.packages("zipcodeR")
#library(zipcodeR)
```
